---
title: "Teaching Survival Git"
subtitle: "Enough Git to Survive as a Data Scientist"
author: "Rob Donald"
date: "`r format(Sys.time(), '%A %d %B %Y')`"

output:
  pdf_document:
    fig_caption: yes
    toc: yes
    toc_depth: '2'
urlcolor: blue
---

```{=tex}
\def\shrug{\texttt{\raisebox{0.75em}{\char`\_}\char`\\\char`\_\kern-0.5ex(\kern-0.25ex\raisebox{0.25ex}
{\rotatebox{45}{\raisebox{-.75ex}"\kern-1.5ex\rotatebox{-90})}}\kern-0.5ex)\kern-0.5ex\char`\_/\raisebox{0.75em}
{\char`\_}}}
```

```{r setup, include=FALSE}
# The fig.pos = 'h' tells LaTeX to put the figure where you ask for it
knitr::opts_chunk$set(echo = TRUE,fig.height=8,fig.width=12,fig.pos = 'h')
options(width=1500)
```

```{r library_setup, include=FALSE}
suppressMessages({suppressWarnings({
    library(dplyr)
    library(tidyr)
    library(readr)
    library(lubridate)
    library(here)
    
    library(ggplot2)
#    library(scales)
    library(ggridges)
    library(RobsRUtils)
    library(pander)
    library(kableExtra)

})})
```


# Introduction

My notes on using git as a data scientist. In other words you *know and accept* you 
should be using git but man is it confusing and almost painful. \small\shrug\normalsize

I've been a software developer and data scientist for more than 30 years and I've used
all sorts of version control systems (started with Apollo DSEE, look that one up). I find
git confusing.


# Moving from SVN to Git


* https://smartbear.com/blog/migrating-from-subversion-to-git-lessons-learned/

* https://dev.to/fpuffer/what-makes-companies-switch-from-svn-to-git-3687

* https://ohshitgit.com/

# Git setup in general

* https://happygitwithr.com/

* https://happygitwithr.com/git-intro.html



# Lightbulb Moments

## Switching back and forth between main and branches

This is from the world famous Jenny Bryan \footnote{
If you don't know who she is look at this link
\url{https://www.rstudio.com/speakers/jenny-bryan/}
and follow her on twitter: @JennyBryan.}

* https://happygitwithr.com/git-branches.html Section 22.2

So you *can* do it and yes git __stash__ is a way but she recommends
the same idea I had read somewhere else about *always* doing a commit
even if it is just with the comment 'WIP xxx'. But she then
goes on to explain how you can get rid of that temporary commit using
terminal (bash) commands like:

```{bash eval=FALSE}
$ git checkout issue-5
$ git reset HEAD^
```


## Merge conflicts

* https://happygitwithr.com/git-branches.html Section 22.4


# Daily routine

```{bash, attr.source='.numberLines', eval=FALSE}
# fetch the changes from the remote
$ git fetch origin

# show commit logs of changes
$ git log master..origin/master

# show diffs of changes
$ git diff master..origin/master

# apply the changes by merge..
$ git merge origin/master

# .. or just pull the changes
$ git pull
```


# How Do I Compare Two PDF Output Files

Well you would think this is easy eh? But the problem is the fact that it is a .pdf file
i.e. __*not*__ a text file. So what you need to do is get the older version's commit SHA1
and then extract that particular commit to somewhere different. Note this command below
is extracting the *whole folder structure* at the point in time from your SHA1.

From this link:

* https://stackoverflow.com/questions/11018411/how-do-i-export-a-specific-commit-with-git-archive


```{bash eval=FALSE}
$ git archive --format zip --output ./CompareResults/Cmp-01.zip 4081d45d
```

Using the above command you have the whole repo as at SHA1 ```4081d45d``` in the zip file ```Cmp-01.zip```

You can of course just pull out a specific file using this command:

```{bash eval=FALSE}
$ git archive --format zip --output ./CompareResults/SingleFile.zip 4081d45d Prospec_Stg_1_N69_ELvsDP.pdf
```


